# Datasets 

This Chapter provides code for datasets produced for the book.

## Log-Returns of International Stock Market Indices Prices {#dt-indices}

### Dataset Location 

./data/global_indices_returns.csv

### Dataset Description 

Log-returns of adjusted prices for the indices identified by the following tickers: ^GSPC, ^FTSE, ^GDAXI, ^N100, ^BVSP

### Original Data Source 

Alpha Vantage

### Code

```{r eval=FALSE, echo=TRUE, cache=TRUE}
library(quantmod)

tickers<-c("^GSPC", "^FTSE", "^GDAXI", "^N100", "^BVSP")

data.env <- new.env()
dataset<- xts() # Only run once


# Download prices from AlphaVantage and calculate log-returns
for(i in 1:length(tickers)) {
  tickers[i]-> symbol
  print(symbol)
  getSymbols(symbol, src="av",
             auto.assign=TRUE,
             output.size="full",
             adjusted=TRUE,
             api.key=config::get()$alpha.vantage.key)

    dataset <- merge(dataset, periodReturn(Ad(get(tickers[i])),period="daily", type='log'))
    rm(symbol)
}

names(dataset)<-tickers

tmp <- tempfile()
write.zoo(dataset,sep=",",file="./data/global_indices_returns.csv")
```

### Dataset Scheme

```{r eval=TRUE, echo=TRUE, cache=TRUE}
head(dataset)
```
