```{r include=FALSE}
library(quantmod)
```


# I/O


## Data Sources

### Alpha Vantage

Alpha Vantage offers free access to pricing data including: 

* Stock Time Series Data; 
* Physical and Digital/Crypto Currencies (e.g., Bitcoin); 
* Technical Indicators and 
* Sector Performances.

The data are available in JSON and CSV format via REST APIs.
The **quantmod** and the **alphavantager** R packages offer a lightweight R interface to the Alpha Vantage API.
For instance, daily stock prices can be obtained with the `quantmod::getSymbols` function as follows:
```{r eval=TRUE, echo=FALSE, cache=TRUE, results='hide', include=FALSE}
getSymbols(Symbols='AAPL', src="av", output.size="full", adjusted=TRUE, api.key=config::get()$alpha.vantage.key)
```
```{r eval=FALSE, echo=TRUE, cache=FALSE}
getSymbols(Symbols='AAPL', src="av", output.size="full", adjusted=TRUE, api.key='your API key')
```
```{r eval=TRUE, echo=TRUE, cache=TRUE}
plot(AAPL$AAPL.Adjusted)
``` 

We called the `quantmod::getSymbols` function with the following arguments:

* `Symbols='AAPL'` defines a character vector specifying the names of each symbol to be loaded, here specified by the symbol of the company Apple Inc.;
* `src="av"` specifies the sourcing method, here defined with the value corresponding to Alpha Vantage;
* `output.size="full"`, strings `compact` and `full` are accepted with the following specifications: `compact` returns only the latest 100 data points; `full` returns the full-length time series of up to 20 years of historical data;
* `adjusted=TRUE`, defines boolean variable to include a column of closing prices adjusted for dividends and splits;
* `api.key`, specifies your Alpha Vantage API key.

### IEX


### Quandl



